---
title: "Day 21 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(igraph)
library(ggplot2)
library(rgl)
library(ggraph)
library(reshape2)
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```


## The Intcode

```{r}
iccomp<-function(prg,inp,starthere,relativebase,quiet){
  i<-starthere
  outp<-c()
  while(prg[i]!=99){
    ## make sure that at least the next three numbers exist
    if((i+3)>length(prg)){
      prg<-c(prg,
             rep(0,(i+3-length(prg))))}
    opcode<-as.character(prg[i]%%100)
    firstp<-(prg[i]%/%100)%%10
    secondp<-(prg[i]%/%1000)%%10
    thirdp<-prg[i]%/%10000
    #   if(quiet==F){cat("i: ",i," fullinst: ",prg[i]," opcode: ",opcode,"\n")}
    switch(opcode,
           ### add
           "1"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             prg[c+1]<-a+b
             i<-i+4},
           ### multiply
           "2"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             prg[c+1]<-a*b
             i<-i+4},
           ### input
           "3"={
             if(firstp==0){a<-prg[i+1]}
             if(firstp==2){a<-prg[i+1]+relativebase}
             if(is.na(a)){a<-0}
             if(length(inp)==0){
               if(quiet==F){cat("waiting---\n")}
               return(list(prg,outp,i,relativebase))
             }else{
               if(quiet==F){cat("input: ",inp[1]," - ")}
               prg[a+1]<-inp[1]
               inp<-inp[-1]
               i<-i+2}},
           ### output
           "4"={
             a<-prg[i+1]
             if(firstp==0){
               # if(quiet==F){cat("looking here: ",a+1,"\n")}
               a<-prg[a+1]}
             if(firstp==2){
               #   if(quiet==F){cat("looking here: ",a+1+relativebase,"\n")}
               a<-prg[a+1+relativebase]}
             if(is.na(a)){a<-0}
             if(quiet==F){cat("output: ",a,"\n")}
             outp<-c(outp,a)
             i<-i+2},
           ### jump if true
           "5"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             if(a!=0){i<-b+1}else{i<-i+3}},
           ### jump if false
           "6"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             if(a==0){i<-b+1}else{i<-i+3}},
           ### less than
           "7"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             if(a<b){prg[c+1]<-1}else{prg[c+1]<-0}
             i<-i+4},
           "8"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             if(a==b){prg[c+1]<-1}else{prg[c+1]<-0}
             i<-i+4},
           "9"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             if(is.na(a)){a<-0}
             relativebase<-relativebase+a
             i<-i+2},
           {cat(i,": unknown opcode",opcode,"\n")
             break})
    prg<-replace(prg,is.na(prg),0)}
  #  cat("ENDING i: ",i," fullinst: ",prg[i]," opcode: ",opcode,"\n")
  return(list(prg,outp,0,relativebase))}
```


```{r}
input<-read_lines("../../AoCData/AOC2019/Day21.txt")
intcode<-as.numeric(unlist(str_split(input,",")))
```

```{r}
inst2ascii<-function(a){
  a<-c(sapply(unlist(str_split(a,"")),utf8ToInt),10)
a}
```


```{r}
ascii2inst<-function(a){
  a<-str_flatten(sapply(a,intToUtf8))
a}
```

I think, jump if (A==0 or if B==0 or if C==0) and D==1

NOT A T (t = 1 if A=0)
OR  T J (j =1 if !A)
NOT B T (t =1 if B=0)
OR  T J (j =1 if B=0)
NOT C T (t =1 if C=0)
OR  T J (j =1 if C=0)
AND D J (if there was a hole earlier, and there's ground, jump)
WALK

----------------
OR A T (t=1 if A=1)
AND B T (t=1 if both A&B are 1)
AND C T (t=1 if A&B&C are 1 - no holes)
NOT T J (J=1 if there's a hole)
AND D J (J=1 if there's a hole and D is ground)


```{r}
x<- c(inst2ascii("OR A T"),
      inst2ascii("AND B T"),
      inst2ascii("AND C T"),
      inst2ascii("NOT T J"),
      inst2ascii("AND D J"),
      inst2ascii("WALK"))
```


```{r}
try<-iccomp(intcode,x,1,0,T)
```

```{r}
whathappened<-try[[2]]
```

```{r}
tail(try[[2]],1)
```
## Part 2

Just a quick "what does it look like when I run" thing:


```{r}
x<- c(inst2ascii("OR A T"),
      inst2ascii("AND B T"),
      inst2ascii("AND C T"),
      inst2ascii("NOT T J"),
      inst2ascii("AND D J"),
      inst2ascii("AND H J"),
      inst2ascii("RUN"))
```



```{r}
try<-iccomp(intcode,x,1,0,T)
```


```{r}
whathappened<-try[[2]]
```

```{r}
cat(ascii2inst(whathappened))
```


Doesn't work because it could make H or E

OR A T (t=1 if A=1)
AND B T (t=1 if both A&B are 1)
AND C T (t=1 if A&B&C are 1 - no holes)
NOT T J (J=1 if there's a hole)
AND J T (clears T)
OR E T 
OR H T  (if there's ground in E or H, T = 1)  (so it can go forward at least one or jump again)
AND T J (maybe jump)
AND D J (J=1 D is ground)

### T = 1 if no holes, T=0 if holes ;

```{r}
x<- c(inst2ascii("OR A T"),
      inst2ascii("AND B T"),
      inst2ascii("AND C T"),
      inst2ascii("NOT T J"),
      inst2ascii("AND J T"),
      inst2ascii("OR E T"),
      inst2ascii("OR H T"),
      inst2ascii("AND T J"),
      inst2ascii("AND D J"),
      inst2ascii("RUN"))
```




try<-iccomp(intcode,x,1,0,F)

```{r}
try<-iccomp(intcode,x,1,0,T)
part2<-tail(try[[2]],1)
part2
```

