---
title: "Day 17 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(ggplot2)
library(reshape2)
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```
## The Intcode

```{r}
iccomp<-function(prg,inp,starthere,relativebase,quiet){
  i<-starthere
  outp<-c()
  while(prg[i]!=99){
    ## make sure that at least the next three numbers exist
    if((i+3)>length(prg)){
      prg<-c(prg,
             rep(0,(i+3-length(prg))))}
    opcode<-as.character(prg[i]%%100)
    firstp<-(prg[i]%/%100)%%10
    secondp<-(prg[i]%/%1000)%%10
    thirdp<-prg[i]%/%10000
    #   if(quiet==F){cat("i: ",i," fullinst: ",prg[i]," opcode: ",opcode,"\n")}
    switch(opcode,
           ### add
           "1"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             prg[c+1]<-a+b
             i<-i+4},
           ### multiply
           "2"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             prg[c+1]<-a*b
             i<-i+4},
           ### input
           "3"={
             if(firstp==0){a<-prg[i+1]}
             if(firstp==2){a<-prg[i+1]+relativebase}
             if(is.na(a)){a<-0}
             if(length(inp)==0){
               if(quiet==F){cat("waiting---\n")}
               return(list(prg,outp,i,relativebase))
             }else{
               if(quiet==F){cat("input: ",inp[1]," - ")}
               prg[a+1]<-inp[1]
               inp<-inp[-1]
               i<-i+2}},
           ### output
           "4"={
             a<-prg[i+1]
             if(firstp==0){
               # if(quiet==F){cat("looking here: ",a+1,"\n")}
               a<-prg[a+1]}
             if(firstp==2){
               #   if(quiet==F){cat("looking here: ",a+1+relativebase,"\n")}
               a<-prg[a+1+relativebase]}
             if(is.na(a)){a<-0}
             if(quiet==F){cat("output: ",a,"\n")}
             outp<-c(outp,a)
             i<-i+2},
           ### jump if true
           "5"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             if(a!=0){i<-b+1}else{i<-i+3}},
           ### jump if false
           "6"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             if(a==0){i<-b+1}else{i<-i+3}},
           ### less than
           "7"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             if(a<b){prg[c+1]<-1}else{prg[c+1]<-0}
             i<-i+4},
           "8"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             if(a==b){prg[c+1]<-1}else{prg[c+1]<-0}
             i<-i+4},
           "9"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             if(is.na(a)){a<-0}
             relativebase<-relativebase+a
             i<-i+2},
           {cat(i,": unknown opcode",opcode,"\n")
             break})
    prg<-replace(prg,is.na(prg),0)}
  #  cat("ENDING i: ",i," fullinst: ",prg[i]," opcode: ",opcode,"\n")
  return(list(prg,outp,0,relativebase))}
```


```{r}
input<-read_lines("../../AoCData/AOC2019/Day17.txt")
intcode<-as.numeric(unlist(str_split(input,",")))

```


## Part 1 
For part 1, just run the intcode - 

```{r}
p1<-iccomp(intcode,c(),1,0,T)
asciiscaffold<-p1[[2]]
```

Having run it, then graph the scaffolding


First, transform the scaffolding into a matrix:
```{r}
nc<-min(which(asciiscaffold==10)-1)
p1asciimatrix<-matrix(ncol=nc,nrow=0)
ln<-1
sl<-c()
for(i in 1:length(asciiscaffold)){
  vw<-asciiscaffold[i]
  ### if at the end of the row
  if(vw==10){
    ### check to see if it's a full row, if so - add it to the matrix
    if(length(sl)==nc){
      p1asciimatrix<-rbind(p1asciimatrix,sl)
      ### otherwise, translate it
      }else{sl<-sapply(sl,intToUtf8)
      cat(sl,"\n")}
    ### move forward a line & reset sl
    ln<-ln+1
    sl<-c()
    ### otherwise, append to the end of the line
  }else{sl<-c(sl,vw)}}
```


Then graph what we have:

```{r}
sdf<-melt(unname(p1asciimatrix))
basicgraph<-ggplot()+
  geom_tile(data=sdf,aes(x=Var2,y=Var1,fill=value))+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position="none")+
  scale_y_reverse()+
  coord_fixed()
basicgraph
```
Kind of pretty - 


To find the intersections - just change from the ascii to 0/1

```{r}
nc<-ncol(p1asciimatrix)
nr<-nrow(p1asciimatrix)
intersectfinder<-p1asciimatrix
intersectfinder[which(p1asciimatrix==46)]<-0
intersectfinder[which(p1asciimatrix==35)]<-1
intersectfinder[which(!p1asciimatrix%in%c(35,46))]<- -1
intersectfinder<-matrix(intersectfinder,ncol=nc,nrow=nr)
```

To find the intersections, add around - if the 3x3 box enclosing the point is sums to 5, it's an intersection.


```{r}
alignparams<-0
for(i in  2:(nrow(intersectfinder)-1)){
  for(j in 2:(ncol(intersectfinder)-1)){
    if(intersectfinder[(i+1),j]+intersectfinder[(i-1),j]+intersectfinder[i,(j+1)]+intersectfinder[i,(j-1)]+intersectfinder[i,j]==5){
      cat(i,j,"\n")
      alignparams<-alignparams+((i-1)*(j-1))}}}
alignparams
```
## Part 2

I'll admit, most of this was done manually - because it was easier:

|   |   |
| ----- | --- | 
| 23,9   | L8 | 
| 23,1   | R12 | 
| 11,1   | R12 | 
| 11,13  | R10 | 
| 21,13  | R10 | 
| 21,3   | R12 | 
| 9,3    | R10 | 
| 9,13   | L8 | 
| 1,13   | R12 | 
| 1,25   | R12 | 
| 13,25  | R10 | 
| 13,15  | R10 | 
| 3,15   | R12 | 
| 3,27   | R10 | 
| 13,27  | L10 | 
| 13,37  | R10 | 
| 23,37  | L6 | 
| 23,43  | L10 | 
| 13,43  | R10 | 
| 13,53  | L6 | 
| 7,53   | R10 | 
| 7,63   | R12 | 
| 19,63  | R10 | 
| 19,53  | L8 | 
| 27,53  | R12 | 
| 27,41  | R12 | 
| 15,41  | R10 | 
| 15,51  | R10 | 
| 25,51  | R12 | 
| 25,39  | R10 | 
| 15,39  | L10 | 
| 15,29  | R10 | 
| 5,29   | L6 | 
| 5,23   |    |




Breaking that down:

```{r}
#A
amess<-"L,8,R,12,R,12,R,10"
#B
bmess<-"L,10,R,10,L,6"
#C
cmess<-"R,10,R,12,R,10"
#MAIN
mainmess<-"A,C,A,C,B,B,C,A,C,B"
```

Concatenate these altogether so the input takes care of itself along the way:

```{r}

mainmess<-c(sapply(unlist(str_split(mainmess,"")),utf8ToInt),10)
amess<-c(sapply(unlist(str_split(amess,"")),utf8ToInt),10)
bmess<-c(sapply(unlist(str_split(bmess,"")),utf8ToInt),10)
cmess<-c(sapply(unlist(str_split(cmess,"")),utf8ToInt),10)

allmess<-c(mainmess,amess,bmess,cmess,utf8ToInt("n"),10)
allmess

```
Then, run it after its set to 2

```{r}
p2prg<-intcode
p2prg[1]<-2
### instead of having doing the input after a prompt it was all front loaded. I didn't get to actual prompts until day 25.
p2out<-iccomp(p2prg,allmess,1,0,T)
```

The answer is 945911

```{r}
finalout<-p2out[[2]]
```

```{r,echo=FALSE,eval=FALSE}
p2asciimatrix<-matrix(ncol=nc,nrow=0)
ln<-1
sl<-c()
for(i in 1:length(finalout)){
  vw<-finalout[i]
  ### if at the end of the row
  if(vw==10){
    ### check to see if it's a full row, if so - add it to the matrix
    if(length(sl)==nc){
      p2asciimatrix<-rbind(p2asciimatrix,sl)
      ### otherwise, translate it
      }else{
        sl<-sapply(sl,intToUtf8)
        cat(str_flatten(sl),"\n")}
    ### move forward a line & reset sl
    ln<-ln+1
    sl<-c()
    ### otherwise, append to the end of the line
  }else{sl<-c(sl,vw)}}
```
```{r,eval=FALSE,echo=FALSE}
sdf2<-melt(unname(p2asciimatrix))
basicgraph<-ggplot()+
  geom_tile(data=sdf2,aes(x=Var2,y=Var1,fill=value))+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position="none")+
  scale_y_reverse()+
  coord_fixed()
basicgraph
```