---
title: "Day 15 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(gganimate)
library(gifski)
library(ggplot2)
library(reshape2)
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```

## The Intcode

```{r}
iccomp<-function(prg,inp,starthere,relativebase,quiet){
#cat("-----------\n")
#cat("-----------\n")
  i<-starthere
  outp<-c()
  while(prg[i]!=99){
    ## make sure that at least the next three numbers exist
    if((i+3)>length(prg)){
      prg<-c(prg,
             rep(0,(i+3-length(prg))))}
    opcode<-as.character(prg[i]%%100)
    firstp<-(prg[i]%/%100)%%10
    secondp<-(prg[i]%/%1000)%%10
    thirdp<-prg[i]%/%10000
 #   if(quiet==F){cat("i: ",i," fullinst: ",prg[i]," opcode: ",opcode,"\n")}
    switch(opcode,
           ### add
           "1"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             prg[c+1]<-a+b
             i<-i+4},
           ### multiply
           "2"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             prg[c+1]<-a*b
             i<-i+4},
           ### input
           "3"={
             if(firstp==0){a<-prg[i+1]}
             if(firstp==2){a<-prg[i+1]+relativebase}
             if(is.na(a)){a<-0}
             if(length(inp)==0){
               if(quiet==F){cat("waiting---\n")}
               return(list(prg,outp,i,relativebase))
             }else{
               if(quiet==F){cat("input: ",inp[1]," - ")}
               prg[a+1]<-inp[1]
           #    if(quiet==F){cat("put it here: ",a+1,"\n")}
               inp<-inp[-1]
               i<-i+2}},
           ### output
           "4"={
             a<-prg[i+1]
            # cat("****\n",a,"******\n")
             if(firstp==0){
           #    if(quiet==F){cat("looking here: ",a+1,"\n")}
               a<-prg[a+1]}
             if(firstp==2){
            #   if(quiet==F){cat("looking here: ",a+1+relativebase,"\n")}
               a<-prg[a+1+relativebase]}
            # cat("****\n",a,"******\n")
             if(is.na(a)){a<-0}
             if(quiet==F){cat("output: ",a,"\n")}
             outp<-c(outp,a)
             i<-i+2},
           ### jump if true
           "5"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             if(a!=0){i<-b+1}else{i<-i+3}},
           ### jump if false
           "6"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             if(a==0){i<-b+1}else{i<-i+3}},
           ### less than
           "7"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             if(a<b){prg[c+1]<-1}else{prg[c+1]<-0}
             i<-i+4},
           "8"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             if(a==b){prg[c+1]<-1}else{prg[c+1]<-0}
             i<-i+4},
           "9"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             if(is.na(a)){a<-0}
             relativebase<-relativebase+a
             i<-i+2},
           {cat(i,": unknown opcode",opcode,"\n")
             break})
    prg<-replace(prg,is.na(prg),0)}
  #  cat("ENDING i: ",i," fullinst: ",prg[i]," opcode: ",opcode,"\n")
  return(list(prg,outp,0,relativebase))}
```


```{r}
input<-read_lines("../../AoCData/AOC2019/Day15.txt")
intcode<-as.numeric(unlist(str_split(input,",")))

```

```{r}
iccomp(intcode,1,1,0,F)[[2]]
iccomp(intcode,2,1,0,F)[[2]]
iccomp(intcode,3,1,0,F)[[2]]
iccomp(intcode,4,1,0,F)[[2]]

```
## Part 1

Going to search this - and it's going to be weird.
The way this works is to test all of the walls as I search so instead of going down & back up a path, just keep going forward  - 

Basic idea is:
1. Get a position from the queue (and the program, starting instruction, and relative base), for each direction:
a. check to see if I already know what is there - if so, no need to bother
b. find out what's there - and store it in my known information for future checks
c. if it is the wall or the oxygen, stop
d. if it's an empty space - put that space (along with the current program, starting instruction, and relative base) in the queue for future exploration.

2. Mark where the oxygen tank is, but keep going to map out everything, just in case

```{r}
findOS<-function(prg){
  beenthere<-dict()
  pq<-priority_queue()
  ### putting in x,y,steps,program,startinst,relativebase
  strt<-list(0,0,0,prg,1,0)
  pq$push(strt,0)
  beenthere$set(str_flatten(c(0,0)," "),c(1,0))
  foundit<-list()
  founditsteps<-c()
  roundcount<-1
  while(pq$size()>0){
    curr<-pq$pop()
    currx<-curr[[1]]
    curry<-curr[[2]]
    currstp<-curr[[3]]
    currprg<-curr[[4]]
    currsi<-curr[[5]]
    currrb<-curr[[6]]
    ### stop and go to the next if at the end
    if(currsi==0){cat("end of the program",currx,curry,"\n")
      next}
    ### for each direction
    ### North
    if(!beenthere$has(str_flatten(c(currx,curry+1)," "))){
      #### find out what's in front of this
      n<-iccomp(currprg,1,currsi,currrb,T)
      #### update been there
      beenthere$set(str_flatten(c(currx,curry+1)," "),c(n[[2]],currstp+1))
      #### if have found the system, add it
      if(n[[2]]==2){
        cat("Foundit! ",currx,curry+1,currstp+1,"\n")
        foundit<-c(foundit,list(c(currx,curry+1)))
        founditsteps<-c(founditsteps,currstp+1)}
      #### otherwise keep going 
      if(n[[2]]==1){
        pq$push(list(currx,curry+1,currstp+1,n[[1]],n[[3]],n[[4]]),currstp+1)}}
    ### South
    if(!beenthere$has(str_flatten(c(currx,curry-1)," "))){
      #### find out what's in front of this
      s<-iccomp(currprg,2,currsi,currrb,T)
      #### update been there
      beenthere$set(str_flatten(c(currx,curry-1)," "),c(s[[2]],currstp+1))
      #### if have found the system, add it
      if(s[[2]]==2){
        cat("Foundit! ",currx,curry-1,currstp+1,"\n")
        foundit<-c(foundit,list(c(currx,curry-1)))
        founditsteps<-c(founditsteps,currstp+1)}
      #### otherwise keep going 
      if(s[[2]]==1){
        pq$push(list(currx,curry-1,currstp+1,s[[1]],s[[3]],s[[4]]),currstp+1)}}
    ### West
    if(!beenthere$has(str_flatten(c(currx-1,curry)," "))){
      #### find out what's in front of this
      w<-iccomp(currprg,3,currsi,currrb,T)
      #### update been there
      beenthere$set(str_flatten(c(currx-1,curry)," "),c(w[[2]],currstp+1))
      #### if have found the system, add it
      if(w[[2]]==2){
        cat("Foundit! ",currx-1,curry,currstp+1,"\n")
        foundit<-c(foundit,list(c(currx-1,curry)))
        founditsteps<-c(founditsteps,currstp+1)}
      #### otherwise keep going
      if(w[[2]]==1){
        pq$push(list(currx-1,curry,currstp+1,w[[1]],w[[3]],w[[4]]),currstp+1)}}
    ### East
    if(!beenthere$has(str_flatten(c(currx+1,curry)," "))){
      #### find out what's in front of this
      e<-iccomp(currprg,4,currsi,currrb,T)
      #### update been there
      beenthere$set(str_flatten(c(currx+1,curry)," "),c(e[[2]],currstp+1))
      #### if have found the system, add it
      if(e[[2]]==2){
        cat("Foundit! ",currx+1,curry,currstp+1,"\n")
        foundit<-c(foundit,list(c(currx+1,curry)))
        founditsteps<-c(founditsteps,currstp+1)}
      #### otherwise keep going 
      if(e[[2]]==1){
        pq$push(list(currx+1,curry,currstp+1,e[[1]],e[[3]],e[[4]]),currstp+1)}}
    roundcount<-roundcount+1
    if(roundcount%%100==0){cat(roundcount,pq$size(),"\n")}
  }
  ### find out where the final step was
  cat(roundcount,"----",paste(foundit),"\n")
  ### list the number of steps, return the "been there"
  list(founditsteps,beenthere)}
```



```{r}
p1<-findOS(intcode)
part1<-p1[[1]]
part1
```
## Part 2

Start by looking at the full map of the area - mapping out everything just in case worked wonders

```{r}
mapme<-p1[[2]]
oxygraph<-as.data.frame(matrix(ncol=4,nrow=0))
coords<-mapme$keys()
fills<-mapme$values()
for(i in 1:length(coords)){
  co<-as.numeric(unlist(str_split(coords[i]," ")))
  co<-list(co[1],co[2],as.character(fills[[i]][1]),fills[[i]][2])
  oxygraph<-rbind(oxygraph,co)}
colnames(oxygraph)<-c("x","y","oxys","minutes")
```


```{r,echo=FALSE}
tlvls<-c("0"="#000088","1"="#FFFFFF","2"="#FF00FF")
ogrstatic<-ggplot()+
  geom_tile(data=oxygraph,aes(x=x,y=y,fill=oxys))+
  geom_point(aes(x=0,y=0),color="#000000")+
  scale_fill_manual(values=tlvls)+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position="none")+
  coord_fixed()

ogr<-ogrstatic+
  transition_states(minutes,wrap=FALSE)+
  shadow_mark()+
  view_follow()

findoxanim<-animate(ogr,nframes=200,renderer = gifski_renderer(),end_pause=15)
ogrstatic
```
Ignore the walls - there's no need to consider them.
```{r}
valspace<-oxygraph%>%rowwise%>%filter(oxys!=0)%>%select(x,y)
```

So then, map out the valid spaces, starting at the watercooler in (12,-12)

```{r}
oxyfill<-function(crdlist){
  checklist<-dict()
  for(i in 1:nrow(crdlist)){
    checklist$set(str_flatten(c(crdlist$x[i],crdlist$y[i])," "),NA)}
  beenthere<-dict()
  pq<-priority_queue()
  ### push x,y,number of steps
  pq$push(c(12,-12,0),0)
  maxsteps<-0
  while(pq$size()>0&&beenthere$size()<checklist$size()){
    curr<-pq$pop()
    currx<-curr[1]
    curry<-curr[2]
    currs<-curr[3]
    if(beenthere$has(str_flatten(c(currx,curry)," "))){next}
    beenthere$set(str_flatten(c(currx,curry)," "),currs)
    maxsteps<-max(maxsteps,currs)
    ### in each direction, check if that's a valid space - if so, put it on the queue. If not - ignore

    ### north
    if(checklist$has(str_flatten(c(currx,curry+1)," "))){
      pq$push(c(currx,curry+1,currs+1),-currs-1)}                           
    ### south
    if(checklist$has(str_flatten(c(currx,curry-1)," "))){
      pq$push(c(currx,curry-1,currs+1),-currs-1)}                           
    ### west
    if(checklist$has(str_flatten(c(currx-1,curry)," "))){
      pq$push(c(currx-1,curry,currs+1),-currs-1)}                           
    ### east
    if(checklist$has(str_flatten(c(currx+1,curry)," "))){
      pq$push(c(currx+1,curry,currs+1),-currs-1)}}
  cat("max steps: ",maxsteps,"\n")
  beenthere}
```

```{r}
p2<-oxyfill(valspace)
```

```{r,eval=FALSE,echo=FALSE}
mapox2<-p1[[2]]
oxyg2<-as.data.frame(matrix(ncol=3,nrow=0))
coords2<-p2$keys()
fills2<-p2$values()
for(i in 1:length(coords2)){
  co<-as.numeric(unlist(str_split(coords2[i]," ")))
  co<-list(co[1],co[2],fills2[[i]])
  oxyg2<-rbind(oxyg2,co)}
colnames(oxyg2)<-c("x","y","mins")
### solely for the animation
oxyg2<-oxyg2%>%rowwise%>%mutate(minutes=mins+340)
```


```{r,eval=FALSE,echo=FALSE}
ogr2<-ogr+
  geom_point(data=oxyg2,aes(x=x,y=y,color=mins))

findoxanim2<-animate(ogr2,nframes=400,renderer = gifski_renderer(),end_pause=15)
findoxanim2
```
```{r,eval=FALSE,echo=FALSE}
anim_save("Day15Part1.gif",findoxanim)
anim_save("Day15Part2.gif",findoxanim2)

```

