---
title: "Day 14 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(igraph)
library(ggplot2)
library(ggraph)
library(reshape2)
library(knitr)
library(dplyr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```

```{r}
input<-read_lines("Day14Sample5.txt")

```

## Part 1
This is recursive.

I think I'm going to do an unholy combination of a about 6 different things... I am building a map so that I can roll up by "distance to ore" & a transform list.
```{r}

rmatrix<-matrix(ncol=4,nrow=0)
for(i in 1:length(input)){
  reaction<-unlist(str_split(input[i], ", | => "))
  ### deal with the result
  result<-tail(reaction,1)
  result<-unlist(str_split(result," "))
  result[[1]]<-as.numeric(result[[1]])
  components<-reaction[-length(reaction)]
  ### make the map
  for(j in 1:length(components)){
    cmp<-components[j]
    cmp<-unlist(str_split(cmp," "))
    cmp[[1]]<-as.numeric(cmp[[1]])
    rmatrix<-rbind(rmatrix,c(result[[1]],result[[2]],cmp[[1]],cmp[[2]]))}}
reactmap<-graph_from_edgelist(rmatrix[,c(4,2)])
rmatrix<-as.data.frame(rmatrix)
colnames(rmatrix)<-c("wcount","wanted","ncount","needed")
rmatrix$wcount<-as.numeric(rmatrix$wcount)
rmatrix$ncount<-as.numeric(rmatrix$ncount)

```



```{r}
ggraph(reactmap,layout="dendrogram")+
  geom_edge_diagonal(aes(start_cap = label_rect(node1.name),
                     end_cap = label_rect(node2.name)), 
                 arrow = arrow(length = unit(2, 'mm')))+
  geom_node_text(aes(label=name))
```





```{r}
ranktable<-as.data.frame(matrix(ncol=2,nrow=0))
for(i in 1:length(V(reactmap))){
  addv<-V(reactmap)[i]$name
  if(addv=="ORE"){fdist<-1
  }else{
    fdist<-max(sapply(all_simple_paths(reactmap,from="ORE",to=addv),
                    function(x){length(names(x))}))}
  ranktable<-rbind(ranktable,list(addv,fdist))
}

colnames(ranktable)<-c("component","rnk")
arrange(ranktable,desc(rnk))
```



Takes in an element and the number wanted and breaks it into the number of components needed to get there
```{r}
breakdown<-function(x,rmat){
  getcomponent<-x$component
  getnumber<-x$cnt
  rmat<-rmat%>%rowwise%>%filter(wanted==getcomponent)
  mplier<-ceiling(getnumber/rmat$wcount[1])
  rmat<-rmat%>%rowwise%>%mutate(ncount=ncount*mplier)%>%
    mutate(component=needed,cnt=ncount)%>%
    select(component,cnt)
  rmat}
```



```{r}
howmuchore<-function(rmat,rtab){
  ### create the starting single fuel
  ###
  inglist<-as.data.frame((matrix(c(component="FUEL",cnt=1),ncol=2,byrow=T)))
  colnames(inglist)<-c("component","cnt")
  inglist$cnt<-as.numeric(inglist$cnt)
  r<-1
  while(any(inglist$component != "ORE")){
    ### find them in the rank table
    inglist<-inglist%>%rowwise%>%
      left_join(rtab,by=join_by(component))%>%
      ### sort the ingredient list by rank (descending)
      arrange(desc(rnk))%>%
      ### remove the rank
      select(-rnk)
    ### grab the top one
    breakme<-breakdown(inglist[1,],rmat)
    ### remove it & break it down into its components
    ### add the new components back in.
    inglist<-rbind(inglist[-1,],breakme)
    ### summarize to keep a list of everything newly needed
    inglist<-inglist%>%group_by(component)%>%summarize(cnt=sum(cnt))
    r<-r+1}
  #### ingredients should be just "ORE"  
  inglist}
```

```{r}
howmuchore(rmatrix,ranktable)
```
## Part 2


```{r}
howmuchore_fuel<-function(rmat,rtab,f){
  ### create the starting single fuel
  ### (see maybe up above)
  inglist<-as.data.frame((matrix(c(component="FUEL",cnt=f),ncol=2,byrow=T)))
  colnames(inglist)<-c("component","cnt")
  inglist$cnt<-as.numeric(inglist$cnt)
  r<-1
  while(any(inglist$component != "ORE")){
    ### find them in the rank table
    inglist<-inglist%>%rowwise%>%
      left_join(rtab,by=join_by(component))%>%
      ### sort the ingredient list by rank (descending)
      arrange(desc(rnk))%>%
      ### remove the rank
      select(-rnk)
    ### grab the top one
    breakme<-breakdown(inglist[1,],rmat)
    ### remove it & break it down into its components
    ### add the new components back in.
    inglist<-rbind(inglist[-1,],breakme)
    ### summarize to keep a list of everything newly needed
    inglist<-inglist%>%group_by(component)%>%summarize(cnt=sum(cnt))
    r<-r+1}
  #### ingredients should be just "ORE"  
  inglist$cnt}
```


Binary Search it - 
is it the elegant way to get this done? no.  But will if *find* the right answer - yes.


```{r}
findatrillion<-function(rmat,rtab){
 possibly<-howmuchore_fuel(rmat,rtab,1)
 cat(possibly,"\n")
 ### set min to be a million over the base, set max to be 10 times that much
 minf<-1000000000000%/%possibly
 maxf<-10*minf
 while(maxf-minf!=1){
   if(maxf-minf==2){attempt<-maxf-1
   }else{attempt<-(maxf+minf)%/%2}
   cat(minf,attempt,maxf,"\n")
   tryit<-howmuchore_fuel(rmat,rtab,attempt)
   if(tryit<=1000000000000){
     minf<-attempt
   }else{maxf<-attempt}}
minf}
```

```{r}
findatrillion(rmatrix,ranktable)
```
