---
title: "Day 19 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(ggplot2)
library(reshape2)
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```

## The Intcode

```{r intcode}
iccomp<-function(prg,inp,starthere,relativebase,quiet){
  i<-starthere
  outp<-c()
  while(prg[i]!=99){
    ## make sure that at least the next three numbers exist
    if((i+3)>length(prg)){
      prg<-c(prg,
             rep(0,(i+3-length(prg))))}
    opcode<-as.character(prg[i]%%100)
    firstp<-(prg[i]%/%100)%%10
    secondp<-(prg[i]%/%1000)%%10
    thirdp<-prg[i]%/%10000
    #   if(quiet==F){cat("i: ",i," fullinst: ",prg[i]," opcode: ",opcode,"\n")}
    switch(opcode,
           ### add
           "1"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             prg[c+1]<-a+b
             i<-i+4},
           ### multiply
           "2"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             prg[c+1]<-a*b
             i<-i+4},
           ### input
           "3"={
             if(firstp==0){a<-prg[i+1]}
             if(firstp==2){a<-prg[i+1]+relativebase}
             if(is.na(a)){a<-0}
             if(length(inp)==0){
               if(quiet==F){cat("waiting---\n")}
               return(list(prg,outp,i,relativebase))
             }else{
               if(quiet==F){cat("input: ",inp[1]," - ")}
               prg[a+1]<-inp[1]
               inp<-inp[-1]
               i<-i+2}},
           ### output
           "4"={
             a<-prg[i+1]
             if(firstp==0){
               # if(quiet==F){cat("looking here: ",a+1,"\n")}
               a<-prg[a+1]}
             if(firstp==2){
               #   if(quiet==F){cat("looking here: ",a+1+relativebase,"\n")}
               a<-prg[a+1+relativebase]}
             if(is.na(a)){a<-0}
             if(quiet==F){cat("output: ",a,"\n")}
             outp<-c(outp,a)
             i<-i+2},
           ### jump if true
           "5"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             if(a!=0){i<-b+1}else{i<-i+3}},
           ### jump if false
           "6"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             if(a==0){i<-b+1}else{i<-i+3}},
           ### less than
           "7"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             if(a<b){prg[c+1]<-1}else{prg[c+1]<-0}
             i<-i+4},
           "8"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             b<-prg[i+2]
             if(secondp==0){b<-prg[b+1]}
             if(secondp==2){b<-prg[b+1+relativebase]}
             if(is.na(a)){a<-0}
             if(is.na(b)){b<-0}
             c<-prg[i+3]
             if(thirdp==2){c<-prg[i+3]+relativebase}
             if(a==b){prg[c+1]<-1}else{prg[c+1]<-0}
             i<-i+4},
           "9"={
             a<-prg[i+1]
             if(firstp==0){a<-prg[a+1]}
             if(firstp==2){a<-prg[a+1+relativebase]}
             if(is.na(a)){a<-0}
             relativebase<-relativebase+a
             i<-i+2},
           {cat(i,": unknown opcode",opcode,"\n")
             break})
    prg<-replace(prg,is.na(prg),0)}
  #  cat("ENDING i: ",i," fullinst: ",prg[i]," opcode: ",opcode,"\n")
  return(list(prg,outp,0,relativebase))}
```


```{r}
input<-read_lines("../../AoCData/AOC2019/Day19.txt")
intcode<-as.numeric(unlist(str_split(input,",")))
```


```{r input}
tractorbeam<-as.data.frame(matrix(ncol=3,nrow=0))
for(i in 0:49){
  for(j in 0:49){
    v<-iccomp(intcode,c(i,j),1,0,T)[[2]]
    tractorbeam<-rbind(tractorbeam,c(i,j,v))}}
colnames(tractorbeam)<-c("x","y","st")
```


```{r graph,echo=FALSE}
 basicgraph<-ggplot()+
  geom_tile(data=tractorbeam,aes(x=x,y=y,fill=st))+
   theme(axis.text.x = element_blank(),
         axis.ticks.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks.y = element_blank(),
         axis.title.y = element_blank(),
         axis.title.x = element_blank(),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         legend.position="none")+
   scale_y_reverse()+
   coord_fixed()
 basicgraph
```


```{r part1}
part1<-sum(tractorbeam$st)
part1
```
## Part 2
I feel like this is a math problem
I have 5 equations,
5 unknowns

Unknowns:
xa,ya ~ (upper right corner of the square)
xb,yb ~ (lower left corner of the square)
c ~ (y intercept of the diagonal of the square that connects points a & b)

ya-xa=c
yb-xb=c
xa-xb=100
ya-yb=100

Then just need the eqs for the upper and lower lines:



```{r findtopline}
upperline<-tractorbeam%>%filter(st==1)%>%filter(y<850)%>%group_by(y)%>%summarise(x=max(x))
upperline
lm(upperline$y~upperline$x)
```
```{r findbottomline}
lowerline<-tractorbeam%>%filter(st==1)%>%group_by(y)%>%summarise(x=min(x))
lowerline
lm(lowerline$y~lowerline$x)
```
ya-xa-c=0
yb-xb-c=0
xa-xb=100
-ya+yb=100
ya-1.0294*xa = -0.5237
yb-1.1397*xb = -0.5624


```{r usesolver}
#Using solve:
#xa, ya, xb, yb, c  
A<-rbind(
  c(1,1,0,0,1),
  c(0,0,1,1,1),
#  c(1,0,-1,0,0),
  c(0,-1,0,1,0),
  c(-1.0294,1,0,0,0),
  c(0,0,-1.1397,1,0))

B<-c(0,0,99,-0.5237,-0.5624)  
solve(A,B)    
```
So the answers are likely close to (1921,1977) & (1822,2076)

Try x from 1772 to 1971 & y from 1927 to 2126 (about 50 up and down)


```{r check_nearby_numbers}
tractor2<-as.data.frame(matrix(ncol=3,nrow=0))
for(i in 1772:1971){
  for(j in 1927:2126){
    v<-iccomp(intcode,c(i,j),1,0,T)[[2]]
    tractor2<-rbind(tractor2,c(i,j,v))}}
colnames(tractor2)<-c("x","y","st")
```


```{r graph2,echo=FALSE}
 part2graph<-ggplot()+
  geom_tile(data=tractor2,aes(x=x,y=y,fill=st))+
  geom_segment(aes(x=1921,y=1977,xend=1822,yend=1977))+
  geom_segment(aes(x=1822,y=1977,xend=1822,yend=2076))+
  geom_segment(aes(x=1822,y=2076,xend=1921,yend=2076))+
  geom_segment(aes(x=1921,y=2076,xend=1921,yend=1977))+
   theme(axis.text.x = element_blank(),
         axis.ticks.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks.y = element_blank(),
         axis.title.y = element_blank(),
         axis.title.x = element_blank(),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         legend.position="none")+
   scale_y_reverse()+
   coord_fixed()
 part2graph
```
This is very close to right - 
Figure out where it gets wide enough
```{r narrowdown}
### filter out where st is 1
findsquare<-tractor2%>%filter(st==1)%>%
  ### for each y, find the highest x
  group_by(y)%>%summarise(x=max(x))%>%
  ### get coordinates for the opposite corner
  rowwise%>%
  mutate(oppx=x-99,oppy=y+99)%>%
  ### find whether the beam is on the opposite corner
  mutate(inbeam=iccomp(intcode,c(x,y),1,0,T)[[2]])%>%
  mutate(oppbeam=iccomp(intcode,c(oppx,oppy),1,0,T)[[2]])%>%
  arrange(y)
findsquare

```


The earliest that the opposite is 1 is the answer

```{r narrowfurther}
findsquare<-findsquare%>%filter(oppbeam==1)%>%arrange(y)
findsquare
```
So, the closestsquare is (1826,1982)  (5 off in each direction!)

```{r part2}
part2<-10000*1826+1982
part2
```

